<i>invivoPKfit</i> and <i>CvTdb</i> figures

### 2022-12-20 Working Rmds/MainAnalysis.Rmd
Add analysis that showed how to impute missing SDs for multi-individual data. Add postprocessing step.

```{r, load_cvt, eval = TRUE}
cvt_data <- fread("../inst/ext/cvt_data_2022_12_13_17_09_45.csv")
```

Add LOQ when available.

PMIDs of documents

```{r}
unique(cvt_data[is.na(conc_time_values.conc),
                .(documents_extraction.pmid,
                  documents_extraction.doi,
                  documents_extraction.other_study_identifier,
                  documents_extraction.url)])
```

For PMID = 1233222, the original document has the LOQ as 0.08 ug/mL.

```{r}
cvt_data[documents_extraction.pmid %In% "1233222",
         c("loq_original",
                                                     "loq_original_units")= .(0.08,
                                                                              "ug/mL")]
```

@@

### do 1-Compartment fit
system.time(
    PK.fit.table.1comp <- fit_all(
      data.set = cvt.data,
      model = "1compartment",
      modelfun = "analytic",
      compound.col = "analyte_name_original",
      dtxsid.col = "dsstox_substance_id",
      cas.col = "dsstox_casrn",
      reference.col = "fk_reference_document_id",
      species.col = "species",
      species.weight.col = "weight_kg",
      species.weight.units.default = "kg",
      dose.col = "dose_level_normalized",
      time.col = "time_hr",
      time.units.default = "hr",
      media.col = "conc_medium_normalized",
      media.units.default = "normalized",
      value.col = "conc",
      units.default = NA,
      route.col = "administration_route_normalized",
      source.col = "document_id",
      loq.col ="calc_loq",
      subject.default = NA,
      info.default = NA,
      suppress.messages = TRUE
  )
)

@@

@@

### do 2-Compartment fit
system.time(
    PK.fit.table.2comp <- fit_all(
      data.set = cvt.data,
      model = "2compartment",
      modelfun = "analytic",
      compound.col = "analyte_name_original",
      dtxsid.col = "dsstox_substance_id",
      cas.col = "dsstox_casrn",
      reference.col = "fk_reference_document_id",
      species.col = "species",
      species.weight.col = "weight_kg",
      species.weight.units.default = "kg",
      dose.col = "dose_level_normalized",
      time.col = "time_hr",
      time.units.default = "hr",
      media.col = "conc_medium_normalized",
      media.units.default = "normalized",
      value.col = "conc",
      units.default = NA,
      route.col = "administration_route_normalized",
      source.col = "document_id",
      loq.col ="calc_loq",
      subject.default = NA,
      info.default = NA,
      suppress.messages = TRUE
    )
)

write.csv(PK.fit.table.2comp, paste("inst/ext/PK_fit_table_2comp_",
                                    timestamp,
                                    ".csv",
                                    sep=""), row.names = FALSE)

@@

system.time(
    PK.fit.table.flat <- fit_all(
      data.set = cvt.data,
      model = "flat",
      modelfun = "analytic",
      compound.col = "analyte_name_original",
      dtxsid.col = "dsstox_substance_id",
      cas.col = "dsstox_casrn",
      reference.col = "fk_reference_document_id",
      species.col = "species",
      species.weight.col = "weight_kg",
      species.weight.units.default = "kg",
      dose.col = "dose_level_normalized",
      time.col = "time_hr",
      time.units.default = "hr",
      media.col = "conc_medium_normalized",
      media.units.default = "normalized",
      value.col = "conc",
      units.default = NA,
      route.col = "administration_route_normalized",
      source.col = "document_id",
      loq.col ="calc_loq",
      subject.default = NA,
      info.default = NA,
      suppress.messages = TRUE
    )
)

@@

```{r}
DF <- preprocess_data(data.set = cvt.data,
                       ratio.data.to.dose = 1,
      compound.col = "analyte_name_original",
      dtxsid.col = "dsstox_substance_id",
      cas.col = "dsstox_casrn",
      reference.col = "fk_reference_document_id",
      species.col = "species",
      species.weight.col = "weight_kg",
      species.weight.units.default = "kg",
      dose.col = "dose_level_normalized",
      time.col = "time_hr",
      time.units.default = "hr",
      media.col = "conc_medium_normalized",
      media.units.default = "normalized",
      value.col = "conc",
      units.default = NA,
      route.col = "administration_route_normalized",
      source.col = "document_id",
      loq.col ="calc_loq",
      subject.default = NA,
      info.default = NA,
      suppress.messages = FALSE)
```

>>>>

```{r, load_cvt, eval = TRUE}
cvt_data <- fread("../inst/ext/cvt_data_2022_12_15_16_20_21.csv")
```

```{r}
cvt <- preprocess_data(data.set = cvt_data,
                       names_list =list(
                       "Compound_Dosed" = "studies.test_substance_name_original",
                              "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
                              "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
                              "Compound_Analyzed" = "series.analyte_name_original",
                              "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
                              "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
                      "Reference" = "documents_reference.id",
                      "Extraction" = "documents_extraction.id",
                      "Species" = "subjects.species",
                      "Weight" ="subjects.weight_kg",
                      "Weight.Units" = NULL,
                      "Dose" = "correct.dose_level_normalized",
                      "Dose.Units" = NULL,
                      "Time" = "conc_time_values.time_hr",
                      "Time.Units" = NULL,
                      "Media" = "series.conc_medium_normalized",
                      "Value" = "conc_time_values.conc",
                      "Value.Units" = NULL,
                      "Route" = "studies.administration_route_normalized",
                      "LOQ" = "series.loq_normalized",
                      "Subject" = "subjects.id",
                              "N_Subjects" = "series.n_subjects_in_series",
                              "Value_SD" = "conc_time_values.conc_sd"),
                    defaults_list =   list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "h",
                      "Value.Units" = "mg/L"),
      ratio_conc_to_dose = 1,
                    calc_loq_factor = 0.45,
                    routes_keep = c("po", "iv"),
                    media_keep = c("blood", "plasma"),
      suppress.messages = FALSE)
```

There are many, many cases where N > 1 but no SD is reported. Can we impute these based on patterns in the cases where SD is reported? Calculate mean and SD for the individual cases:

```{r}
cvt <- as.data.table(cvt)
cvt_summarized <- cvt[N_Subjects==1,
                      .(meanval = mean(Value, na.rm = TRUE),
                        sdvals = sd(Value, na.rm = TRUE),
                        n = .N),
                      by = .(Reference, DTXSID, Media, Dose)]
ggplot(cvt_summarized) +
  geom_point(aes(x = meanval, y = sdvals, color = n)) +
  scale_x_log10() + scale_y_log10() + scale_color_viridis_c(trans = "log10")

summary(lm(log10(sdvals) ~ log10(meanval), data = cvt_summarized))
```

The relationship between sample SD and sample mean seems to be almost the identity line.

What about the multi-subject cases with reported SD?

```{r}
ggplot(cvt[N_Subjects > 1],
       aes(x = Value, y = Value_SD)) +
  geom_point(aes(x = Value, y = Value_SD, color = N_Subjects))  +
  scale_x_log10() + scale_y_log10() + geom_smooth(method = "lm")

summary(lm(log10(Value_SD) ~ log10(Value), data = cvt[N_Subjects > 1]))
```

This one isn't quite the identity line but still, we can get some idea.

These ones with entirely missing SDs are different from the studies where only a few SDs are missing. When only a few are missing, it was usually because the data were pulled from a plot where the error bars were smaller thant the plotting symbol on a few points. Instead of recording the size of the symbol as the error bars, the error was recorded as NA. But when all SDs are missing, it is usually because the study did not report the SDs in the first place.

In this case, it seems like a reasonable choice to impute the mean value itself as the SD. This relation seems to be borne out by the individual-anumal data, and it is more conservative than the relation suggested by the (limited) multi-animal data. I will make that change.

@@

### do 1-Compartment fit
system.time(
    PK.fit.table.1comp <- fit_all(
      data.set = cvt_data,
      model = "1compartment",
      modelfun = "analytic",
      names_list =list(
                      "Compound_Dosed" = "studies.test_substance_name_original",
                      "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
                      "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
                      "Compound_Analyzed" = "series.analyte_name_original",
                      "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
                      "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
                      "Reference" = "documents_reference.id",
                      "Extraction" = "documents_extraction.id",
                      "Species" = "subjects.species",
                      "Weight" ="subjects.weight_kg",
                      "Weight.Units" = NULL,
                      "Dose" = "studies.dose_level_normalized",
                      "Dose.Units" = NULL,
                      "Time" = "conc_time_values.time_hr",
                      "Time.Units" = NULL,
                      "Media" = "series.conc_medium_normalized",
                      "Value" = "conc_time_values.conc",
                      "Value.Units" = NULL,
                      "Route" = "studies.administration_route_normalized",
                      "LOQ" = "series.loq",
                      "Subject" = "subjects.id",
                      "N_Subjects" = "series.n_subjects_in_series",
                      "Value_SD" = "conc_time_values.conc_sd"),
                    defaults_list =   list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "h",
                      "Value.Units" = "mg/L"),
      ratio_conc_to_dose = 1,
                    calc_loq_factor = 0.45,
                    routes_keep = c("po", "iv"),
                    media_keep = c("blood", "plasma"),
      optimx_args = list(
                      "method" = "bobyqa",
                      "itnmax" = 1e6,
                      "control" = list("maximize" = TRUE,
                                       "kkt" = FALSE)
                    ),
      suppress.messages = TRUE
  )
)

@@

post_1comp <- postprocess_data(PK_fit = PK.fit.table.1comp,
                               model = "1compartment")

write.csv(post_1comp,
          paste("inst/ext/PK_fit_postprocess_1comp_",
                                    timestamp,
                                    ".csv",
                                    sep=""),
          row.names = FALSE)

@@

### do 2-Compartment fit
system.time(
    PK.fit.table.2comp <- fit_all(
      data.set = cvt_data,
      model = "flat",
      modelfun = "analytic",
      names_list =list(
                      "Compound_Dosed" = "studies.test_substance_name_original",
                      "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
                      "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
                      "Compound_Analyzed" = "series.analyte_name_original",
                      "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
                      "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
                      "Reference" = "documents_reference.id",
                      "Extraction" = "documents_extraction.id",
                      "Species" = "subjects.species",
                      "Weight" ="subjects.weight_kg",
                      "Weight.Units" = NULL,
                      "Dose" = "studies.dose_level_normalized",
                      "Dose.Units" = NULL,
                      "Time" = "conc_time_values.time_hr",
                      "Time.Units" = NULL,
                      "Media" = "series.conc_medium_normalized",
                      "Value" = "conc_time_values.conc",
                      "Value.Units" = NULL,
                      "Route" = "studies.administration_route_normalized",
                      "LOQ" = "series.loq",
                      "Subject" = "subjects.id",
                      "N_Subjects" = "series.n_subjects_in_series",
                      "Value_SD" = "conc_time_values.conc_sd"),
                    defaults_list =   list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "h",
                      "Value.Units" = "mg/L"),
      ratio_conc_to_dose = 1,
                    calc_loq_factor = 0.45,
                    routes_keep = c("po", "iv"),
                    media_keep = c("blood", "plasma"),
      optimx_args = list(
                      "method" = "bobyqa",
                      "itnmax" = 1e6,
                      "control" = list("maximize" = TRUE,
                                       "kkt" = FALSE)
                    ),
      suppress.messages = TRUE
    )
)

write.csv(PK.fit.table.2comp, paste("inst/ext/PK_fit_table_2comp_",
                                    timestamp,
                                    ".csv",
                                    sep=""), row.names = FALSE)
post_2comp <- postprocess_data(PK_fit = PK.fit.table.2comp,
                               model = "2compartment")

write.csv(post_2comp,
          paste("inst/ext/PK_fit_postprocess_2comp_",
                                    timestamp,
                                    ".csv",
                                    sep=""),
          row.names = FALSE)
```

@@

system.time(
    PK.fit.table.flat <- fit_all(
     data.set = cvt_data,
      model = "flat",
      modelfun = "analytic",
      names_list =list(
                      "Compound_Dosed" = "studies.test_substance_name_original",
                      "DTXSID_Dosed" = "chemicals_dosed.dsstox_substance_id",
                      "CAS_Dosed" = "chemicals_dosed.dsstox_casrn",
                      "Compound_Analyzed" = "series.analyte_name_original",
                      "DTXSID_Analyzed" = "chemicals_analyzed.dsstox_substance_id",
                      "CAS_Analyzed" = "chemicals_analyzed.dsstox_casrn",
                      "Reference" = "documents_reference.id",
                      "Extraction" = "documents_extraction.id",
                      "Species" = "subjects.species",
                      "Weight" ="subjects.weight_kg",
                      "Weight.Units" = NULL,
                      "Dose" = "studies.dose_level_normalized",
                      "Dose.Units" = NULL,
                      "Time" = "conc_time_values.time_hr",
                      "Time.Units" = NULL,
                      "Media" = "series.conc_medium_normalized",
                      "Value" = "conc_time_values.conc",
                      "Value.Units" = NULL,
                      "Route" = "studies.administration_route_normalized",
                      "LOQ" = "series.loq",
                      "Subject" = "subjects.id",
                      "N_Subjects" = "series.n_subjects_in_series",
                      "Value_SD" = "conc_time_values.conc_sd"),
                    defaults_list =   list(
                      "Weight.Units" = "kg",
                      "Dose.Units" = "mg/kg",
                      "Time.Units" = "h",
                      "Value.Units" = "mg/L"),
      ratio_conc_to_dose = 1,
                    calc_loq_factor = 0.45,
                    routes_keep = c("po", "iv"),
                    media_keep = c("blood", "plasma"),
      optimx_args = list(
                      "method" = "bobyqa",
                      "itnmax" = 1e6,
                      "control" = list("maximize" = TRUE,
                                       "kkt" = FALSE)
                    ),
      suppress.messages = TRUE
    )
)

@@
